---
output: 
  pdf_document:
    latex_engine: xelatex
params:
  stop_id: 2488
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, error = FALSE)
```

\begin{huge}
\textbf{This bus stop is closing June 11}
\end{huge}

Starting June 11, Metro's bus network will change a lot. Buses will be faster and there will fewer transfers for most. But this stop will be closing permanently.

```{r warning=FALSE}
library(tidyverse)
library(sf)
library(tmap)
library(tidytransit)
library(OpenStreetMap)

# read in closed stops
closed_stops <- read_csv("busstop-closedlist.csv") |>
  rename(stop_id = `Stop ID`,
         direction = Direction,
         stop_name = `Stop Name`)

#geocode stops with GTFS files
gtfs <- read_gtfs("metro-2023-04-17-gtfs.zip")

stops <- gtfs$stops

# doing an inner join because the closure list has separate IDs for each direction of a stop
# but the GTFS file does not
# closure list also has a bunch of stops that currently don't have service
closed_stops_sf <- closed_stops |>
  inner_join(stops, by = c("stop_id" = "stop_code")) |>
  st_as_sf(coords = c("stop_lon", "stop_lat"))

# next we need all the stops that will still have service after the redesign
open_stops <- stops |>
  anti_join(closed_stops, by = c("stop_code" = "stop_id")) |>
  st_as_sf(coords = c("stop_lon", "stop_lat"))
st_crs(open_stops) <- 4326
st_crs(closed_stops_sf) <- 4326

#filter to the stop defined in the input parameter

closed_stop <- closed_stops_sf |> 
  filter(stop_id == params$stop_id)

#create 1km buffer around closed stop
closed_stop_buffered <- closed_stop |> 
  st_buffer(dist = 1000)

x <- open_stops |> 
  st_filter(closed_stop_buffered, .predicate = st_intersects)

stop_desc <- paste0(closed_stop_buffered$stop_name.x, 
                   " (",
                   closed_stop_buffered$stop_id,
                   ").")
```

This is stop `r stop_desc`

The map below shows the nearest stops that will still be open. Find out more at mymetrobus.com/redesign or call 608 266 4466.

```{r}

library(tmaptools)
osm <- tmaptools::read_osm(bb(st_union(x, closed_stop)), ext=1.1)
tmap_mode("plot")
tm_shape(osm) +
tm_rgb(alpha = .4) +
  tm_shape(closed_stop_buffered) +
  tm_dots("darkgrey", size = 2) +
  tm_shape(closed_stop) +
  tm_symbols(shape = 4, size = 3) +
  tm_shape(x) +
  tm_dots("darkgrey", size = 1, alpha = .7)
```

```{r}
x |> 
  rowwise() |> 
  mutate(dist = st_distance(geometry, closed_stop),
         dist_mi = units::set_units(dist, mile)) |> 
  select(stop_name, stop_id, dist_mi) |> 
  st_drop_geometry() |> 
  arrange(dist_mi) |> 
  head(10) |> 
  knitr::kable()

```


This sign brought to you by Friends of Metro.
